Voice Classification by Neural Networks

### 2015-05-16 paper.tex
Started stochastic gradient descendescent & ROC sections; minor edits

\documentclass{article}

@@

A curious development in macine learning was the \emph{neural network}, a biologically inspired machine learning model. Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons. We shall soon offer a mathematically rigorous definition of the neural network.

@@

% insert stuff about gradient descent!
Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (for some number of parameters $w_i$, which we write as a vector), find the parameters such that we minimize some \emph{objective function} dependent on both data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$. For our purposes, we will minimize mean-square error, defined as
\beq
\label{eq:meansqerror}
E(\vec{w}) = \sum_k E_k(\vec{w}) = \sum_k (y_k - f(\vec{w})(x_k))^2
\eeq

In practice, there will be numerous parameters $w_i$ (each $\alpha$ and $\beta$ in \cref{eq:hiddenlayer,eq:netvalue} is a $w_i$), so finding a global minimum will be challenging: setting the derivatives of $E$ to 0 and solving the resulting nonlinear equations (our $f$ will involve the nonlinear logistic $\sigma$ and softmax functions) is unfeasible. Instead, we will employ a technique known as gradient descent.

The basic idea is simple. Suppose we are looking for low points in some terrain. We pick an arbitrary location to begin. Then, we walk downhill (in the direction with the steepest drop) a certain distance. We stop, then check which way will take us down the steepest path from our new location. We walk downhill a certain distance again, then repeat the process. Clearly, gradient descent will \emph{not} necessarily find the global minimum, but rather a local one.

We formalize this idea as follows. Given an objective function $E$, we start at some arbitrary $\vec{w}_0$. To find a better estimate for $\vec{w}$, we iteratively go ``downhill" from $\vec{w}_{n-1}$ to $\vec{w}_{n}$, computing
\beq
\vec{w}_n = \vec{w}_{n-1} - \eta\vec{\nabla} E(\vec{w}_{n-1})
\eeq
where $\eta > 0$ is some fixed step size (the ``certain distance" in ou
r hill analogy). The gradient $\vec{\nabla} E$ represents the direction of steepest ascent of the function $E$. Gradient descent is extremely useful for our purposes because for the $E$ we are using, the resulting derivatives $\frac{\partial E}{\partial w_i}$ are relatively simple to determine.

@@

% TODO stochastic gradient descent

@@

% TODO
% ROC curves
% AUC metric

>>>>

\documentclass[10pt]{article}

@@

A curious development in macine learning was the \emph{neural network} (or informally, a neural net), a biologically inspired machine learning model. Indeed, we see how the nodes in \cref{fig:neuralnet} resemble neurons sending information to other neurons. We shall soon offer a mathematically rigorous definition of the neural network.

@@

% insert stuff about gradient descent!
Formally, we can rephrase our problem as follows: Given a set of $K$ data points $(x_k, y_k)$ and a family of functions $f(\vec{w})$ (where we write the parameters as components of vector $\vec{w}$), find the parameters that minimize some \emph{objective function} dependent on both the data points and the parameters that represents in some way the inaccuracy of the estimates $f(\vec{w})(x_k)$ of $y_k$. For our purposes, we will minimize mean-square error, defined as
\beq
\label{eq:meansqerror}
E(\vec{w}) = \sum_k E_k(\vec{w}) = \sum_k (y_k - f(\vec{w})(x_k))^2
\eeq

In practice, there will be numerous parameters (each $\alpha$ and $\beta$ in \cref{eq:hiddenlayer,eq:netvalue} is a component of $\vec{w}$), so finding a global minimum will be challenging: setting the derivatives of $E$ to 0 and solving the resulting nonlinear equations (our $f$ will involve the nonlinear logistic $\sigma$ and softmax functions) is unfeasible. Instead, we will employ a technique known as gradient descent.

The basic idea is simple. Suppose we are looking for low points in some terrain. We pick an arbitrary location to begin. Then, we walk downhill (in the direction of steepest descent) a certain distance. We stop, then check which way will take us down the steepest path from our new location. We walk downhill in that direction a certain distance again, then repeat the process. Clearly, gradient descent will \emph{not} necessarily find the global minimum, but rather a local one.

We formalize this idea as follows. Given an objective function $E$, we start with some arbitrary set of weights $\vec{w}_0$. To find a better estimate for $\vec{w}$, we iteratively go ``downhill" from $\vec{w}_{n-1}$ to $\vec{w}_{n}$, computing
\beq
\label{eq:graddescent}
\vec{w}_n = \vec{w}_{n-1} - \eta\vec{\nabla} E(\vec{w}_{n-1})
\eeq
where $\eta > 0$ is some fixed step size (the ``certain distance" in our hill analogy). The gradient $\vec{\nabla} E$ represents the direction of steepest ascent of the function $E$. Gradient descent is extremely useful for our purposes because for the $E$ we are using, as the resulting derivatives $\frac{\partial E}{\partial w}$ are relatively simple to determine.

@@

\subsubsection{Stochastic Gradient Descent}

Returning to \cref{eq:meansqerror}, we find that our error can be expressed as the sum of the errors on each individual data point. Instead of minimizing $E = \sum_k E_k$, we can try to minimize an \emph{individual} $E_k$ (imagine replacing $E$ with $E_k$ in \cref{eq:graddescent}). In the long run, we have reason to expect this procedure also minimizes $E$.

% TODO continue

@@

Though our goal will ultimately be to find the feature vectors and training procedure that will result in the \emph{best} model, we have not yet introduced any metric for evaluating our models. We shall now do just that.

% ROC space (false pos, true pos)
% ROC curve
% AUC
